### 2017-03-06 paper/paper.Rmd
org files

---
title: "Final Project Paper"
output:
  html_document: default
  html_notebook: default
---

#### Isabella Spaletta, Marisa Nanakul, Christopher Ross, Kai Yoshino
##### Affiliation: University of Washington - Information School

### Abstract

### Introduction

Currently, there are many different sites and organizations that collect data on cities or large metropolitan areas. The Federal Bureau of Investigation (FBI) collects crime data every year for individual cities within the United States, and the Bureau of Economic Analysis (BEA) gathered an overall metropolitan area Gross Domestic Product (GDP) every year since 2010. There are a few different organizations or articles that have looked at a comparison between crime rates and the GDP of specific metropolitan areas, but our specific analysis does not consider outside covariates when comparing crime rate and GDP of metropolitan areas. Some of the related work we will discuss takes the change in economy after the 2008 financial crisis into account. One other research group took into account economic uncertainty whereas we will not be taking into account outside covariates and comparing crime rate and GDP at face value. We want to look at the GDP and crime relationship unaffected by other covariates to see if there is a correlation between the two variables.

**NEED KEY FINDINGS**

Our analysis focuses on violent and property crime impact on city GDPs without covariates. The analysis is a unique contribution because rather than analyzing from the state level, we are choosing to focus on the city level, adding an extra area of granularity. In terms of covariates, we are not accounting for confounding covariates such as uncertainty demonstrated by Goulas. We are observing and comparing violent and property crime trends to city GDP trends.

### Related Work

In the past, most research has been focused around states in the U.S. and countries around the world. In “Falling Together” [1], data from the U.S. Department of Justice from 2007-2010 was used to analyze the relationship between property crime rates and GDP per person in all 50 states. The results found that property crime rates and GDP both decreased during this time frame. Specifically after the recession, the general trend was as GDP decreased so did property crimes. The cause of the trend is unknown.

“The Puzzling Relationship Between Crime and the Economy” [2] delves into the different conclusions economists and criminologists conclude about the relationship between crime and the economy. Criminologists believe that tough economic times increase the likelihood of a person committing a crime. They also believe that economic anxieties cause people to consume more mind-altering substances, leading to increased violence. Economists believe better economic conditions increase the likelihood of a person committing a crime. With technology becoming more visible and ingrained into people’s lives, criminals are able to see what’s possible to steal. Analysis on violent and property crimes from 1960-2012 suggests that there is no relationship between GDP and crime, and that crime is episodic.

In “Economic growth and crime: does uncertainty matter?” [3] 25 countries (including the U.S.) are analyzed with a macroeconomic uncertainty metric to observe whether there is a relationship between crime and GDP. The researchers found that if GDP uncertainty is negative, the crime uncertainty is positive, meaning that crime uncertainty negatively impacts GDP uncertainty.

### Data Prep

The preparation of our data, specifically combining the crime and GDP datasets, was a very important and large aspect of conducting this research. The data given to us was very different in formatting and was therefore fairly difficult to combine. The FBI crime dataset had city names in one column rather than metropolitan areas (some cities combined), and in another column the state names were fully written out instead of having state abbreviations. All of the crime datasets were also separated by year, whereas the GDP dataset had columns of GDP data separated by year.

To begin combining these two datasets, we joined all of the crime datasets together so that every year was accounted for in each city within the dataset. From here, we decided that since the GDP dataset had less cities/metropolitan areas to account for, we would take all of the corresponding cities from the crime dataset and add the GDP for each of those cities to the crime dataset. We then removed any city within the crime dataset that did not include a GDP (was not included in the GDP dataset). New rows were added for the metropolitan areas that included more than one city (e.g. Seattle-Tacoma-Bellevue) and added up the crimes in each of the cities to get a total for the new row added. This is important because to keep the GDP accurate, we had to keep the cities combined as a metropolitan area, but we could account for the crimes by adding the counts together and doing analysis from the new totals. We also removed any metropolitan area from the GDP dataset that included multiple states.

Our final dataset came out as a final CSV that included any city or metropolitan areas that were included in both the crime and GDP dataset. For each of these cities or metropolitan areas, we included the crime data for 2010-2015 and the GDP for 2010-2015. This came out to include _____**(how many rows did we end up with for our data?)** cities or metropolitan areas.

One thing to note is that to reproduce this research you will have to use the CSV we created or any other researcher could find other data sources for GDP and crime for the years 2010 to 2015 and create their own new dataset.

### Analysis

### Results

### Discussion

### Future Work

### References

Goulas, Eleftherios, and Athina Zervoyianni. "Economic Growth and Crime: Does Uncertainty Matter?" Applied Economics Letters 20.5 (2013): 420-27. Web.

Roman, John. "The Puzzling Relationship Between Crime and the Economy." CityLab. The Atlantic, 23 Sept. 2013. Web. 05 Mar. 2017.

The Economist Online. "Falling Together." The Economist. The Economist Newspaper, 26 Sept. 2011. Web. 05 Mar. 2017.

### 2017-03-06 resource/index.Rmd
fix data

---
title: "Understanding the Affects of GDP and Population on Crime in US Cities"
output: html_notebook
---
&nbsp;

#### Isabella Spaletta, Marisa Nanakul, Christopher Ross, Kai Yoshino
##### Affiliation: University of Washington - Information School
&nbsp;

### Purpose

### 2017-03-06 resource/resources.R
add resource file

@@ -1,0 +1,57 @@
+library(dplyr)
+library(plotly)
+library(ggplot2)
+library(tidyr)
+
+# Read in data sets
+cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
+gdp <- read.csv("data/csv/gdp_crime.csv")
+
+# convert all NA values to 0, ya not great but it makes things work
+gdp[is.na(gdp)] <- 0
+
+gdp$gdp_per_person_2010 <- gdp$gdp_2010 * 1000000 / gdp$Population_2010
+
+# Add total crime for each year
+gdp <- gdp %>% 
+        mutate(total_crime_2010 = (Violent_Crime_2010 + Property_Crime_2010), 
+               total_crime_2011 = (Violent_Crime_2011 + Property_Crime_2011),
+               total_crime_2012 = (Violent_Crime_2012 + Property_Crime_2012),
+               total_crime_2013 = (Violent_Crime_2013 + Property_Crime_2013),
+               total_crime_2014 = (Violent_Crime_2014 + Property_Crime_2014),
+               total_crime_2015 = (Violent_Crime_2015 + Property_Crime_2015))
+
+gdp <- gdp %>% 
+      mutate(crime_per_person_2010 = (total_crime_2010 / Population_2010),
+             crime_per_person_2011 = (total_crime_2011 / Population_2011),
+             crime_per_person_2012 = (total_crime_2012 / Population_2012),
+             crime_per_person_2013 = (total_crime_2013 / Population_2013),
+             crime_per_person_2014 = (total_crime_2014 / Population_2014),
+             crime_per_person_2015 = (total_crime_2015 / Population_2015))
+
+gdp <- gdp %>% 
+  mutate(crime_change_10_11 = (crime_per_person_2011 - crime_per_person_2010),
+         crime_change_11_12 = (crime_per_person_2012 - crime_per_person_2011),
+         crime_change_12_13 = (crime_per_person_2013 - crime_per_person_2012),
+         crime_change_14_13 = (crime_per_person_2014 - crime_per_person_2013),
+         crime_change_15_14 = (crime_per_person_2015 - crime_per_person_2014))
+
+
+
+# Poisson regression of gdp per person and population vs crime per person. 
+poisson.2010 <- glm(crime_per_person_2010 ~ gdp_per_person_2010 + Population_2010, data = gdp, 
+    family = poisson(link = "log"))
+
[…]

### 2017-03-07 resource/index.Rmd
making long format for gdp and crime

@@
 ---
-title: "Understanding the Affects of GDP and Population on Crime in US Cities"
-output: html_notebook
+title: "Understanding GDP relations to Crime in US Cities"
+author: "Kai Yoshina, Isabella Spaletta, Chrisopther Ross, Marisa Nanakul"
+date: "March 6, 2017"
+output: html_document
 ---
-&nbsp;
 
-#### Isabella Spaletta, Marisa Nanakul, Christopher Ross, Kai Yoshino
-##### Affiliation: University of Washington - Information School
-&nbsp;
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+source("resources.R")
+```
 
-### Purpose
+### What is the data?
+
+The data that we used in this anaylsis is combination of a data set that has crime in US cities over a 5 year time persion (2010-2015) and the GDP for cities in the same time period.

### 2017-03-07 make-long-gdp.R

@@ -1,0 +1,43 @@
+library(dplyr)
+
+gdp <- read.csv("data/csv/gdp_crime.csv")
+
+gdp[is.na(gdp)] <- 0
+
+gdp$gdp_per_person_2010 <- gdp$gdp_2010 * 1000000 / gdp$Population_2010
+gdp$gdp_per_person_2011 <- gdp$gdp_2011 * 1000000 / gdp$Population_2011
+gdp$gdp_per_person_2012 <- gdp$gdp_2012 * 1000000 / gdp$Population_2012
+gdp$gdp_per_person_2013 <- gdp$gdp_2013 * 1000000 / gdp$Population_2013
+gdp$gdp_per_person_2014 <- gdp$gdp_2014 * 1000000 / gdp$Population_2014
+gdp$gdp_per_person_2015 <- gdp$gdp_2015 * 1000000 / gdp$Population_2015
+
+# Add total crime for each year
+gdp <- gdp %>% 
+  mutate(total_crime_2010 = (Violent_Crime_2010 + Property_Crime_2010), 
+         total_crime_2011 = (Violent_Crime_2011 + Property_Crime_2011),
+         total_crime_2012 = (Violent_Crime_2012 + Property_Crime_2012),
+         total_crime_2013 = (Violent_Crime_2013 + Property_Crime_2013),
+         total_crime_2014 = (Violent_Crime_2014 + Property_Crime_2014),
+         total_crime_2015 = (Violent_Crime_2015 + Property_Crime_2015))
+
+gdp <- gdp %>% 
+  mutate(crime_per_person_2010 = (total_crime_2010 / Population_2010),
+         crime_per_person_2011 = (total_crime_2011 / Population_2011),
+         crime_per_person_2012 = (total_crime_2012 / Population_2012),
+         crime_per_person_2013 = (total_crime_2013 / Population_2013),
+         crime_per_person_2014 = (total_crime_2014 / Population_2014),
+         crime_per_person_2015 = (total_crime_2015 / Population_2015))
+
+gdp <- gdp %>% 
+  mutate(crime_change_10_11 = (crime_per_person_2011 - crime_per_person_2010),
+         crime_change_11_12 = (crime_per_person_2012 - crime_per_person_2011),
+         crime_change_12_13 = (crime_per_person_2013 - crime_per_person_2012),
+         crime_change_14_13 = (crime_per_person_2014 - crime_per_person_2013),
[…]

### 2017-03-07 resource/resources.R

@@ -3,14 +3,23 @@ library(plotly)
 library(ggplot2)
 library(tidyr)
 
-# Read in data sets
+# Read in data sets while in file
 cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
 gdp <- read.csv("data/csv/gdp_crime.csv")
+
+# read in data sets while in RMD file 
+# cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
+# gdp <- read.csv("../data/csv/gdp_crime.csv")
 
 # convert all NA values to 0, ya not great but it makes things work
 gdp[is.na(gdp)] <- 0
 
 gdp$gdp_per_person_2010 <- gdp$gdp_2010 * 1000000 / gdp$Population_2010
+gdp$gdp_per_person_2011 <- gdp$gdp_2011 * 1000000 / gdp$Population_2011
+gdp$gdp_per_person_2012 <- gdp$gdp_2012 * 1000000 / gdp$Population_2012
+gdp$gdp_per_person_2013 <- gdp$gdp_2013 * 1000000 / gdp$Population_2013
+gdp$gdp_per_person_2014 <- gdp$gdp_2014 * 1000000 / gdp$Population_2014
+gdp$gdp_per_person_2015 <- gdp$gdp_2015 * 1000000 / gdp$Population_2015
 
 # Add total crime for each year
 gdp <- gdp %>% 
@@ -45,13 +54,35 @@ poisson.2010 <- glm(crime_per_person_2010 ~ gdp_per_person_2010 + Population_201
 summary(poisson.2010)
 # f(x) = -3.082 + 0.0000009385(GDP) + 0.0000001263(POP)
 
+poisson.2011 <- glm(crime_per_person_2011 ~ gdp_per_person_2011 + Population_2011, data = gdp, 
+                    family = poisson(link = "log"))
+
+summary(poisson.2011)
+
+poisson.2012 <- glm(crime_per_person_2012 ~ gdp_per_person_2012 + Population_2012, data = gdp, 
+                    family = poisson(link = "log"))
+
+summary(poisson.2012)
+
+poisson.2013 <- glm(crime_per_person_2013 ~ gdp_per_person_2013 + Population_2013, data = gdp, 
+                    family = poisson(link = "log"))
+
+summary(poisson.2013)
+
+poisson.2014 <- glm(crime_per_person_2014 ~ gdp_per_person_2014 + Population_2014, data = gdp, 
[…]

### 2017-03-07 paper/paper.Rmd
added future work

@@
 
+While we looked at GDP and crime rate, to extend the work we have done here, other researchers can also compare GDP and crime rate per city with other covariates that may also affect these two.
+One example is other related work looks at economic uncertainty as a covariate and there are other covariates that could possibly show a correlation between the GDP and crime rate of a city such as education, population density, health metrics, and many more.
+
+One other aspect to consider is the particular datasets used and how the data is measured.
+Since we are doing by city, crime rate to us is defined as total number of crimes divided by the population of each city or metropolitan area within the dataset.
+The GDP dataset that we chose to use was in a very unfortunate format when trying to combine it with the crime dataset, so I would suggest that in future work an alternative GDP dataset is used that is by actual city in the US rather than metropolitan area.
+
 ### References

### 2017-03-07 resource/index.Rmd
add new gdp data and melt stuff

@@
 
-The data that we used in this anaylsis is combination of a data set that has crime in US cities over a 5 year time persion (2010-2015) and the GDP for cities in the same time period.
+The data that we used in this anaylsis is a combination of a data set that has crime in US cities over a 5 year time period (2010-2015) and the GDP for cities in the same time period.

### 2017-03-07 resource/resources.R

@@ -2,10 +2,12 @@ library(dplyr)
 library(plotly)
 library(ggplot2)
 library(tidyr)
+library(reshape2)
 
 # Read in data sets while in file
 cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
 gdp <- read.csv("data/csv/gdp_crime.csv")
+long.gdp <- read.csv("for-long.csv")
 
 # read in data sets while in RMD file 
 # cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
@@ -79,10 +81,21 @@ poisson.2015 <- glm(crime_per_person_2015 ~ gdp_per_person_2015 + Population_201
 
 summary(poisson.2015)
 
-# Remove rows with zeros
-# row_sub = apply(dd, 1, function(row) all(row !=0 ))
-##Subset as usual
-# dd[row_sub,]
+total <- cities.all %>% 
+          filter(grepl("-", City))
+columbia <- cities.all %>% 
+            filter(City == "Columbia")
+
+test <- melt(gdp, id.vars = c("State", "City", "id"))
+
+population <- test %>% 
+            filter(grepl("Population_", variable))
+
+total.crime <- test %>% 
+            filter(grepl("total_crime_", variable))
+
+total.gdp <- test %>% 
+            filter(grepl("gdp_per_person", variable))
 
 
 

### 2017-03-07 resource/resources.R
change column names

@@ -97,5 +97,6 @@ total.crime <- test %>%
 total.gdp <- test %>% 
             filter(grepl("gdp_per_person", variable))
 
-
+crime.per.person <- test %>% 
+            filter(grepl(""))
 

### 2017-03-07 resource/index.Rmd
added to resource

--- 2017-03-06 resource/index.Rmd
@@
 ---
-title: "Understanding the Affects of GDP and Population on Crime in US Cities"
-output: html_notebook
+title: "Understanding GDP relations to Crime in US Cities"
+author: "Kai Yoshino, Isabella Spaletta, Chrisopther Ross, Marisa Nanakul"
+date: "March 6, 2017"
+output: html_document
 ---
-&nbsp;
 
-#### Isabella Spaletta, Marisa Nanakul, Christopher Ross, Kai Yoshino
-##### Affiliation: University of Washington - Information School
-&nbsp;
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+library(knitr)
+source("resources.R")
+```
 
-### Purpose
+### What Are The Datasets?
+
+The two datasets that we used in this anaylsis are one for crime based data on U.S.
+cities over a 5 year time period (2010-2015) and one for the GDP for metropolitan areas in the U.S.
+over a 5 year time period (2010-2015).
+
+#### FBI Crime Dataset
+
+Below is an example of what the crime dataset looks like (shows all 5 years):
+
+```{r, echo=FALSE}
+  kable(cities.all[1:5,])
+```
+
+##### Observations Made About the Dataset
+* The size of the dataset is {rows, columns}:
+
+```{r, echo=FALSE}
+  dim(cities.all)
+```
+
+* Column Names and Data Types of the crime table:
+
+```{r, echo=FALSE}
+  colnames(cities.all)
+```
+
+```{r, echo=FALSE}
+  sapply(cities.all, typeof)
+```
+
+##### Other Observations & Changes Made to Dataset
+* All crime datasets were individually captured by year, but we combined them to make one large crime dataset.
+* If a city has a crime value of NA, this means this data could not be found, not that any of these crimes did not occur. NA does not equal 0 in this dataset.
+* Violent crimes is defined of four offenses: murder and nonnegligent manslaughter, rape, robbery, and aggravated assault.
+* Property crimes is defined of four offenses: burglary, larceny theft, motor vehicle theft, and arson.
+<br>
+<br>
+
+#### GDP Dataset
+
+```{r, echo=FALSE}
+  gdp[1:5,] %>% 
+  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015") %>%
+  kable()
+```
+
+##### Observations Made About the Dataset
+* The size of the dataset is {rows, columns}:
+
+```{r, echo=FALSE}
+  dim(gdp %>% 
+  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
+```
+
+* Column Names and Data Types of the crime table:
+
+```{r, echo=FALSE}
+  colnames(gdp %>% 
+  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
+```
+
+```{r, echo=FALSE}
+  sapply(gdp %>% 
+  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"), typeof)
+```
+
+##### Other Observations & Changes Made to Dataset
+* All of this GDP data was added to the entire crime dataset.
+* We included the GDP each of the five years and GDP per person for each of the five years.
+<br>
+<br>

### 2017-03-07 resource/resources.R

@@ -4,12 +4,12 @@ library(ggplot2)
 library(tidyr)
 
 # Read in data sets while in file
-cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
-gdp <- read.csv("data/csv/gdp_crime.csv")
+# cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
+# gdp <- read.csv("data/csv/gdp_crime.csv")
 
 # read in data sets while in RMD file 
-# cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
-# gdp <- read.csv("../data/csv/gdp_crime.csv")
+cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
+gdp <- read.csv("../data/csv/gdp_crime.csv")
 
 # convert all NA values to 0, ya not great but it makes things work
 gdp[is.na(gdp)] <- 0

### 2017-03-07 resource/resources.R
resource.r changed

@@ -16,21 +16,21 @@ long.gdp <- read.csv("for-long.csv")
 # convert all NA values to 0, ya not great but it makes things work
 gdp[is.na(gdp)] <- 0
 
-gdp$gdp_per_person_2010 <- gdp$gdp_2010 * 1000000 / gdp$Population_2010
-gdp$gdp_per_person_2011 <- gdp$gdp_2011 * 1000000 / gdp$Population_2011
-gdp$gdp_per_person_2012 <- gdp$gdp_2012 * 1000000 / gdp$Population_2012
-gdp$gdp_per_person_2013 <- gdp$gdp_2013 * 1000000 / gdp$Population_2013
-gdp$gdp_per_person_2014 <- gdp$gdp_2014 * 1000000 / gdp$Population_2014
-gdp$gdp_per_person_2015 <- gdp$gdp_2015 * 1000000 / gdp$Population_2015
+gdp$gdp_per_person_2010 <- gdp$X2010_gdp * 1000000 / gdp$Population_2010
+gdp$gdp_per_person_2011 <- gdp$X2011_gdp * 1000000 / gdp$Population_2011
+gdp$gdp_per_person_2012 <- gdp$X2012_gdp * 1000000 / gdp$Population_2012
+gdp$gdp_per_person_2013 <- gdp$X2013_gdp * 1000000 / gdp$Population_2013
+gdp$gdp_per_person_2014 <- gdp$X2014_gdp * 1000000 / gdp$Population_2014
+gdp$gdp_per_person_2015 <- gdp$X2015_gdp * 1000000 / gdp$Population_2015
 
 # Add total crime for each year
 gdp <- gdp %>% 
-        mutate(total_crime_2010 = (Violent_Crime_2010 + Property_Crime_2010), 
-               total_crime_2011 = (Violent_Crime_2011 + Property_Crime_2011),
-               total_crime_2012 = (Violent_Crime_2012 + Property_Crime_2012),
-               total_crime_2013 = (Violent_Crime_2013 + Property_Crime_2013),
-               total_crime_2014 = (Violent_Crime_2014 + Property_Crime_2014),
-               total_crime_2015 = (Violent_Crime_2015 + Property_Crime_2015))
+        mutate(total_crime_2010 = (Violent_crime_2010 + Property_crime_2010), 
+               total_crime_2011 = (Violent_crime_2011 + Property_crime_2011),
+               total_crime_2012 = (Violent_crime_2012 + Property_crime_2012),
[…]

### 2017-03-08 resource/resources.R
regressions added

@@ -7,13 +7,13 @@ library(reshape2)
 # Read in data sets while in file
 
 cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
-#gdp <- read.csv("data/csv/gdp_crime.csv")
-#long.gdp <- read.csv("for-long.csv")
+gdp <- read.csv("data/csv/gdp_crime.csv")
+long.gdp <- read.csv("for-long.csv")
 
 
 # read in data sets while in RMD file 
-cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
-gdp <- read.csv("../data/csv/gdp_crime.csv")
+# cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
+# gdp <- read.csv("../data/csv/gdp_crime.csv")
 
 # convert all NA values to 0, ya not great but it makes things work
 gdp[is.na(gdp)] <- 0
@@ -96,9 +96,35 @@ population <- test %>%
 total.crime <- test %>% 
             filter(grepl("total_crime_", variable))
 
-total.gdp <- test %>% 
+person.gdp <- test %>% 
             filter(grepl("gdp_per_person", variable))
 
 crime.per.person <- test %>% 
-            filter(grepl(""))
+            filter(grepl("crime_per_person", variable))
 
+# Adding prefix to ready for join 
+colnames(population)[5] <- paste("population", colnames(population)[5], sep = "_")
+colnames(total.crime)[5] <- paste("total_crime", colnames(total.crime)[5], sep = "_")
+colnames(person.gdp)[5] <- paste("person_gdp", colnames(person.gdp)[5], sep = "_")
+colnames(crime.per.person)[5] <- paste("person_crime", colnames(crime.per.person)[5], sep = "_")
+
+time.poisson.data <- population
+colnames(person.gdp)[5]
+
+time.poisson.data$population_value <- population$population_value
+time.poisson.data$total_crime_value <- total.crime$total_crime_value
+time.poisson.data$person_gdp_value <- person.gdp$person_gdp_value
+time.poisson.data$person_crime_value <- crime.per.person$person_crime_value
+
+
+# This is a poisson regression looks at the crime committed per people 
+poisson.per.person.time <- glm(person_crime_value ~ population_value + person_gdp_value, data = time.poisson.data,
+                               family = poisson(link = "log"))
+
+summary(poisson.per.person.time)
+
[…]

### 2017-03-08 paper/paper.Rmd
added to resource
added to paper

@@
 
+********** Marisa
+The key findings of our research showed that a poisson regression showed no correlation between the GDP per person and crime per person in each metropolitan area with population as another independent variable.
+Between all of the years from 2010 to 2015, there was no significance between GDP per person, crime per person, and population of a metropolitan area for any compared years.
+From this, our contribution to society and information in general is that just looking at a metropolitan area's economy and population is not a good determining factor on the number of crimes permitted by person.
+
 ### Introduction
@@
 We want to look at the GDP and crime relationship unaffected by other covariates to see if there is a correlation between the two variables.
-
-**NEED KEY FINDINGS**
 
@@
 For each of these cities or metropolitan areas, we included the crime data for 2010-2015 and the GDP for 2010-2015.
-This came out to include _____**(how many rows did we end up with for our data?)** cities or metropolitan areas.
+This came out to include 366 cities or metropolitan areas.
+We also added columns such as crime per person in each metropolitan area and gdp per person in each metropolitan area to run our analyses on.
 
 One thing to note is that to reproduce this research you will have to use the CSV we created or any other researcher could find other data sources for GDP and crime for the years 2010 to 2015 and create their own new dataset.
+
+Another thing to note is that 
+'violent crimes' is defined by four offenses: murder and nonnegligent manslaughter, rape, robbery, and aggravated assault. 
+'property crimes' is defined by four offenses: burglary, larceny theft, motor vehicle theft, and arson.
 
@@
 
-Goulas, Eleftherios, and Athina Zervoyianni. "Economic Growth and Crime: Does Uncertainty Matter?" Applied Economics Letters 20.5 (2013): 420-27. Web.
+[3] Goulas, Eleftherios, and Athina Zervoyianni. "Economic Growth and Crime: Does Uncertainty Matter?" Applied Economics Letters 20.5 (2013): 420-27. Web.
 
-Roman, John. "The Puzzling Relationship Between Crime and the Economy." CityLab. The Atlantic, 23 Sept. 2013. Web. 05 Mar. 2017.
+[2] Roman, John. "The Puzzling Relationship Between Crime and the Economy." CityLab. The Atlantic, 23 Sept. 2013. Web. 05 Mar. 2017.
 
-The Economist Online. "Falling Together." The Economist. The Economist Newspaper, 26 Sept. 2011. Web. 05 Mar. 2017.
+[1] The Economist Online. "Falling Together." The Economist. The Economist Newspaper, 26 Sept. 2011. Web. 05 Mar. 2017.

### 2017-03-08 resource/index.Rmd
added to paper

@@
 ---
-title: "Understanding GDP relations to Crime in US Cities"
+title: "Understanding GDP Relations to Crime in US Cities"
 author: "Kai Yoshino, Isabella Spaletta, Chrisopther Ross, Marisa Nanakul"

### 2017-03-08 paper/paper.Rmd
updated intro

@@
 Currently, there are many different sites and organizations that collect data on cities or large metropolitan areas.
-The Federal Bureau of Investigation (FBI) collects crime data every year for individual cities within the United States, and the Bureau of Economic Analysis (BEA) gathered an overall metropolitan area Gross Domestic Product (GDP) every year since 2010.
-There are a few different organizations or articles that have looked at a comparison between crime rates and the GDP of specific metropolitan areas, but our specific analysis does not consider outside covariates when comparing crime rate and GDP of metropolitan areas.
+The Federal Bureau of Investigation (FBI) collects crime data and population every year for individual cities within the United States and the Bureau of Economic Analysis (BEA) gathered an overall metropolitan area Gross Domestic Product (GDP) every year since 2010.
+There are a few different organizations or articles that have looked at a comparison between crime rates and the GDP of specific metropolitan areas, but our specific analysis considers population as another independent variable when comparing crime rate and GDP of metropolitan areas.
 Some of the related work we will discuss takes the change in economy after the 2008 financial crisis into account.
-One other research group took into account economic uncertainty whereas we will not be taking into account outside covariates and comparing crime rate and GDP at face value.
-We want to look at the GDP and crime relationship unaffected by other covariates to see if there is a correlation between the two variables.
+One other research group took into account economic uncertainty whereas we will focus more on the count data of a metropolitan area (population, GDP).
+We want to look at the GDP and crime relationship affected by population to see if GDP per person and population of a metropolitan area have any sort of impact on the crimes per person committed within that metropolitan area.
 
-Our analysis focuses on violent and property crime impact on city GDPs without covariates.
+Our analysis focuses on violent and property crime impact on city GDPs with population as another independent variable.
 The analysis is a unique contribution because rather than analyzing from the state level, we are choosing to focus on the city level, adding an extra area of granularity.
-In terms of covariates, we are not accounting for confounding covariates such as uncertainty demonstrated by Goulas.
-We are observing and comparing violent and property crime trends to city GDP trends.
+In terms of covariates, we are accounting for the population change between years along with the changes in GDP and crime.
+We are observing and comparing violent and property crime trends to metropolitan area GDP and population trends.
 

### 2017-03-08 resource/index.Rmd
worked on RMD

--- 2017-03-07 resource/index.Rmd
@@
 ```{r, echo=FALSE}
-  colnames(cities.all)
-```
-
-```{r, echo=FALSE}
   sapply(cities.all, typeof)
@@
   gdp[1:5,] %>% 
-  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015") %>%
+  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015") %>%
   kable()
@@
   dim(gdp %>% 
-  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
+  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
 ```
@@
   colnames(gdp %>% 
-  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
+  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
 ```
@@
   sapply(gdp %>% 
-  select_("gdp_2010", "gdp_2011", "gdp_2012", "gdp_2013", "gdp_2014", "gdp_2015", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"), typeof)
+  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"), typeof)
 ```
@@
 <br>
+
+##### What all this means? 
+The main purpose of combining these two data sets was to determine if there was some sort of correlation between a cities Gross Domestic Product and the amount of crime that happens within it.
+
+``` {r echo=FALSE}
+
+```

### 2017-03-08 resource/crime-specific.R

@@ -6,14 +6,14 @@ library(reshape2)
 
 # Read in data sets while in file
 
-# cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
-# gdp <- read.csv("data/csv/gdp_crime.csv")
-# long.gdp <- read.csv("for-long.csv")
+cities.all <- read.csv("data/csv/Cities-2010-2015.csv") 
+gdp <- read.csv("data/csv/gdp_crime.csv")
+long.gdp <- read.csv("for-long.csv")
 
 
 # read in data sets while in RMD file 
-cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
-gdp <- read.csv("../data/csv/gdp_crime.csv")
+# cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
+# gdp <- read.csv("../data/csv/gdp_crime.csv")
 
 # convert all NA values to 0, ya not great but it makes things work
 gdp[is.na(gdp)] <- 0
@@ -27,20 +27,20 @@ gdp$gdp_per_person_2015 <- gdp$X2015_gdp * 1000000 / gdp$Population_2015
 
 # Add total crime for each year
 gdp <- gdp %>% 
-        mutate(total_crime_2010 = (Violent_crime_2010 + Property_crime_2010), 
-               total_crime_2011 = (Violent_crime_2011 + Property_crime_2011),
-               total_crime_2012 = (Violent_crime_2012 + Property_crime_2012),
-               total_crime_2013 = (Violent_crime_2013 + Property_crime_2013),
-               total_crime_2014 = (Violent_crime_2014 + Property_crime_2014),
-               total_crime_2015 = (Violent_crime_2015 + Property_crime_2015))
+  mutate(total_crime_2010 = (Violent_crime_2010 + Property_crime_2010), 
+         total_crime_2011 = (Violent_crime_2011 + Property_crime_2011),
+         total_crime_2012 = (Violent_crime_2012 + Property_crime_2012),
+         total_crime_2013 = (Violent_crime_2013 + Property_crime_2013),
+         total_crime_2014 = (Violent_crime_2014 + Property_crime_2014),
+         total_crime_2015 = (Violent_crime_2015 + Property_crime_2015))
 
 gdp <- gdp %>% 
-      mutate(crime_per_person_2010 = (total_crime_2010 / Population_2010),
-             crime_per_person_2011 = (total_crime_2011 / Population_2011),
[…]

### 2017-03-08 paper/paper.Rmd

@@
 The FBI crime dataset had city names in one column rather than metropolitan areas (some cities combined), and in another column the state names were fully written out instead of having state abbreviations.
-All of the crime datasets were also separated by year, whereas the GDP dataset had columns of GDP data separated by year.
+All of the crime datasets were separated by year equalling six different datasets, whereas the GDP dataset had columns of GDP data separated by year equalling only one dataset that all had to be combined.
 
-To begin combining these two datasets, we joined all of the crime datasets together so that every year was accounted for in each city within the dataset.
+To begin combining these two datasets, we joined all of the crime datasets together so that every year was accounted for in each city within the dataset along with the population of each city.
 From here, we decided that since the GDP dataset had less cities/metropolitan areas to account for, we would take all of the corresponding cities from the crime dataset and add the GDP for each of those cities to the crime dataset.
@@
 New rows were added for the metropolitan areas that included more than one city (e.g.
-Seattle-Tacoma-Bellevue) and added up the crimes in each of the cities to get a total for the new row added.
-This is important because to keep the GDP accurate, we had to keep the cities combined as a metropolitan area, but we could account for the crimes by adding the counts together and doing analysis from the new totals.
-We also removed any metropolitan area from the GDP dataset that included multiple states.
+Seattle-Tacoma-Bellevue) and added up the population and crimes in each of the cities to get a total for the new row added.
+This is important because to keep the GDP accurate, we had to keep the cities combined as a metropolitan area, but we could account for the population and crimes by adding the counts together and doing analysis from the new totals.
+We also removed any metropolitan area from the GDP dataset that included multiple states and had to change any value in the data set that was NA to 0 to be able to combine these datasets properly.
 
 Our final dataset came out as a final CSV that included any city or metropolitan areas that were included in both the crime and GDP dataset.
-For each of these cities or metropolitan areas, we included the crime data for 2010-2015 and the GDP for 2010-2015.
+For each of these cities or metropolitan areas, we included the population of each metropolitan area, crime data for 2010-2015, and the GDP for 2010-2015.
 This came out to include 366 cities or metropolitan areas.
@@
 One thing to note is that to reproduce this research you will have to use the CSV we created or any other researcher could find other data sources for GDP and crime for the years 2010 to 2015 and create their own new dataset.
+You must also change any NA in the datset to 0 to combine the data.
 
@@
 
-While we looked at GDP and crime rate, to extend the work we have done here, other researchers can also compare GDP and crime rate per city with other covariates that may also affect these two.
-One example is other related work looks at economic uncertainty as a covariate and there are other covariates that could possibly show a correlation between the GDP and crime rate of a city such as education, population density, health metrics, and many more.
+While we looked at GDP and crime rate, to extend the work we have done here, other researchers can also compare GDP and crime rate per city with other independent variables other than population that may also affect the correlation.
+One example is other related work looks at economic uncertainty as an independten variable and there are other covariates that could possibly show a correlation between the GDP and crime rate of a city such as education, population density, health metrics, and many more.
 
 One other aspect to consider is the particular datasets used and how the data is measured.
-Since we are doing by city, crime rate to us is defined as total number of crimes divided by the population of each city or metropolitan area within the dataset.
+Since we are doing by metropolitan area, crime rate to us is defined as total number of crimes divided by the population of each city or metropolitan area within the dataset.
 The GDP dataset that we chose to use was in a very unfortunate format when trying to combine it with the crime dataset, so I would suggest that in future work an alternative GDP dataset is used that is by actual city in the US rather than metropolitan area.

### 2017-03-08 resource/resources.R
pulling

@@ -3,6 +3,7 @@ library(plotly)
 library(ggplot2)
 library(tidyr)
 library(reshape2)
+library(data.table)
 
 # Read in data sets while in file
 
@@ -127,4 +128,98 @@ poisson.pop.crime <- glm(person_crime_value ~ population_value, data = time.pois
                                family = poisson(link = "log"))
 summary(poisson.pop.crime)
 
-p
+gdp_rate_change <- gdp
+
+# adds columns: by year changes in gdp per person  
+gdp_rate_change$gdp_10_11 <- with(gdp_rate_change, (X2011_gdp - X2010_gdp) / mean(c(Population_2010, Population_2011)))
+gdp_rate_change$gdp_11_12 <- with(gdp_rate_change, (X2012_gdp - X2011_gdp) / mean(c(Population_2011, Population_2012)))
+gdp_rate_change$gdp_12_13 <- with(gdp_rate_change, (X2013_gdp - X2012_gdp) / mean(c(Population_2012, Population_2013)))
+gdp_rate_change$gdp_13_14 <- with(gdp_rate_change, (X2014_gdp - X2013_gdp) / mean(c(Population_2013, Population_2014)))
+gdp_rate_change$gdp_14_15 <- with(gdp_rate_change, (X2015_gdp - X2014_gdp) / mean(c(Population_2014, Population_2015)))
+
+# adds columns: by year changes in population 
+gdp_rate_change$Population_10_11 <- with(gdp_rate_change, (Population_2011 - Population_2010) / mean(c(Population_2010, Population_2011)))
+gdp_rate_change$Population_11_12 <- with(gdp_rate_change, (Population_2012 - Population_2011) / mean(c(Population_2010, Population_2011)))
+gdp_rate_change$Population_12_13 <- with(gdp_rate_change, (Population_2013 - Population_2012) / mean(c(Population_2010, Population_2011)))
+gdp_rate_change$Population_13_14 <- with(gdp_rate_change, (Population_2014 - Population_2013) / mean(c(Population_2010, Population_2011)))
+gdp_rate_change$Population_14_15 <- with(gdp_rate_change, (Population_2015 - Population_2014) / mean(c(Population_2010, Population_2011)))
+
+# adds columns: by year changes in total crime per person
+gdp_rate_change$total_crime_10_11 <- with(gdp_rate_change, (total_crime_2011 - total_crime_2010) / mean(c(Population_2010, Population_2011)))
[…]

### 2017-03-08 resource/resources.R
rate change linear regressions

@@ -128,7 +128,6 @@ poisson.pop.crime <- glm(person_crime_value ~ population_value, data = time.pois
                                family = poisson(link = "log"))
 summary(poisson.pop.crime)
 
-<<<<<<< HEAD
 gdp_rate_change <- gdp
 
 # adds columns: by year changes in gdp per person  

### 2017-03-08 resource/index.Rmd
random cleanup

@@
 library(knitr)
+library(rmarkdown)
 source("resources.R")
@@
 ### What Are The Datasets?
-
+()
 The two datasets that we used in this anaylsis are one for crime based data on U.S.

### 2017-03-08 resource/resources.R

@@ -13,8 +13,8 @@ library(data.table)
 
 
 # read in data sets while in RMD file 
-cities.all <- read.csv("../data/csv/Cities-2010-2015.csv")
-gdp <- read.csv("../data/csv/gdp_crime.csv")
+cities.all <- read.csv("data/csv/Cities-2010-2015.csv")
+gdp <- read.csv("data/csv/gdp_crime.csv")
 
 # convert all NA values to 0, ya not great but it makes things work
 gdp[is.na(gdp)] <- 0
@@ -197,15 +197,11 @@ colnames(property.crime.rate.change)[5] <- paste("property_crime_rate_change", c
 
 linear.change.data <- population.change
 
-
 linear.change.data$population_change_value <- population.change$population_change_value
 linear.change.data$total_crime_rate_change_value <- total.crime.rate.change$total_crime_rate_change_value
 linear.change.data$gdp_rate_change_value <- gdp.rate.change$gdp_rate_change_value
 linear.change.data$violent_crime_rate_change_value <- violent.crime.rate.change$violent_crime_rate_change_value
 linear.change.data$property_crime_rate_change_value <- property.crime.rate.change$property_crime_rate_change_value
-
-
-sapply(linear.change.data, class)
 
 # use this to remove zero values - change tmp back to linear.change.data if you want to use full data frame
 tmp <- setDT(linear.change.data)[, .SD[!any(.SD[, -1, with = F] == 0)], by = City]

### 2017-03-08 resource/index.Rmd
worked on resource

@@
 
-##### Other Observations & Changes Made to Dataset
-* All of this GDP data was added to the entire crime dataset.
-* We included the GDP each of the five years and GDP per person for each of the five years.
+### What We Found
+The main purpose of combining these two data sets was to determine if there was some sort of correlation between a cities Gross Domestic Product, their population and the amount of crime that happens within it.
+Now when initially looking at a data set like this it is very easy to just used the hard numbers to determine if there is any correlation between these three things.
+If for example you were to look at just the population of a city vs the number of crimes committed in that city you would find that, why yes they are related.
+And if you were to plot this data it would look something like this:
+
+_Note: The population data has been passed through a *log* function which basically compresses the data so that data that was closer is easily visible and those data points that were outliers are now in a closer grouping._
+
+``` {r echo=FALSE, message=FALSE, warning=FALSE}
+total.crime.vs.pop <- ggplot(gdp, aes(x=log(Population_2015), y=total_crime_2015, color="red")) +
+                    geom_point() + 
+                    geom_smooth(col = "black") +
+                    ggtitle("Population and Total Crimes in 2015") +
+                    labs(x = "Log(Population)", y= "Total Crimes Committed")
+
+total.crime.vs.pop
+```
+
+While it makes sense that as Populuation increases the number of crimes increase, but we believed that just that information is not what should be the focus when looking at trends in crime.
+Instead we created a value that we named *Crime Rate* that is the total crimes committed in a city divided by the population.
+This new value is the total number of crimes per person within a city, or what we refer to as the crime rate.
+_Note:_ This is not to be confused with the more common term "crime rate" that refers to the number of crimes committed per 100,000 people.
 <br> 
 <br> 
+So what we did is we took the newly made variable *Crime Rate* and compared it with the population of the different cities and their Gross Domestic Product per Person.
+If we break all that down here are how the different data points relate to each other
+``` {r echo=FALSE, message=FALSE, warning=FALSE}
+colnames(time.poisson.data)[4] <- "Year"
 
-##### What all this means? 
-The main purpose of combining these two data sets was to determine if there was some sort of correlation between a cities Gross Domestic Product and the amount of crime that happens within it.
+time.poisson.data$Year <- gsub("Population_", "", time.poisson.data$Year)
 
-``` {r echo=FALSE}
+crime.vs.pop.2015 <- ggplot(time.poisson.data, aes(x=population_value, y=person_crime_value, color=Year)) +
+                  geom_point(shape=1) +
+                  stat_smooth(method = "glm", col = "black") +
+                  ggtitle("Population and Crimes commited per person over 2010-2015") +
+                  labs(x = "Population", y= "Crime Rates")
+
+gdp.vs.crime <- ggplot(time.poisson.data, aes(x=person_gdp_value, y=person_crime_value, color=Year)) +
+                  geom_point(shape=1) +
+                  stat_smooth(method = "glm", col = "black") +
+                  ggtitle("Crimes per Person and GDP per person over 2010-2015") +
+                  labs(x = "GDP per Person", y= "Crime Rate")
+
+gdp.vs.pop.2015 <- ggplot(time.poisson.data, aes(x=population_value, y=person_gdp_value, color=Year)) +
+                  geom_point(shape=1) +
+                  stat_smooth(method = "glm", col = "black") +
+                  ggtitle("Population and GDP per person over 2010-2015") +
+                  labs(x ="Population", y="GDP Per Person")
+
+gdp.vs.pop.2015
 
 ```
+
+```{r echo=FALSE, warning=FALSE, message=FALSE}
+crime.vs.pop.2015
+```
+
+```{r echo=FALSE, warning=FALSE, message=FALSE}
+gdp.vs.crime
+```
+
+Each of the graphs has been given a regression line to make it easier to see the different trends that occur within each of the comparissons.
+Yet to truly understand if there is a statistical signifigance to each of these we needed to run a multiple variable Poisson Regression.
+A Poisson Regression allows us to work with count data and data that is not distributed normally.
+Meaning when you look at the graph it doesn't look like a bell.
+As the graphs demonstrate this is very much true for our data set.
+After running our Poisson Regression we came up with this equation:
+
+#### Crime Rate = 4.938e-7(GDP Per Person) - 5.72e-8(Population) - 3.086
+
+What this breaks down to is that for every one dollar that GDP per person increase the Crime rate increase by 0.0000004938 and for every person added to the population crme rate decreases by 0.0000000527.
+Theses numbers seem incredibly small but when you are talking about millions of dollars in GDP and thousands of people it becomes increasingly important.
+
+###

### 2017-03-08 resource/index.Rmd
pulling

--- 2017-03-08 resource/index.Rmd
@@
 ```
+
+``` {r echo=FALSE}
+linear.pop.gdp.crime <- lm(total_crime_rate_change_value ~ population_change_value + gdp_rate_change_value, data = linear.change.data)
+summary(linear.pop.gdp.crime)
+
+plot(lm(total_crime_rate_change_value ~ population_change_value, data = linear.change.data))
+
+plot(lm(total_crime_rate_change_value ~ population_change_value, data = linear.change.data))
+```

### 2017-03-08 resource/index.Rmd
more work on resource

--- 2017-03-08 resource/index.Rmd
@@
 ### What Are The Datasets?
-()
+
 The two datasets that we used in this anaylsis are one for crime based data on U.S.
@@
 <br>
+
+###Crime Rate with GDP and Population
+
 So what we did is we took the newly made variable *Crime Rate* and compared it with the population of the different cities and their Gross Domestic Product per Person.
@@
 Theses numbers seem incredibly small but when you are talking about millions of dollars in GDP and thousands of people it becomes increasingly important.
+After computing this equation it was still found to not be statistically signifigant, meaning that the GDP per Person and the Population had no real impact on the Crime Rate.
 
-###
+### Crime Rate Change
+
+After looking at just the crime rate over the 5 year time period we created a new variable which was the difference in crime rate, GDP per person, and Population change between the years.
+At this point we had started to work more with countinous data than count data, so we decided to use a standard linear regression to find the equation and signifigance of the data.
+
+``` {r echo=FALSE, message=FALSE, warnings=FALSE}
+rate.crime.vs.pop <- ggplot(tmp, aes(x=population_change_value, y=total_crime_rate_change_value)) +
+                  geom_point(shape=1) +
+                  stat_smooth(method = "lm", col = "red") +
+                  ggtitle("Population and Total Crime Rate Change") +
+                  labs(x ="Population", y="Total Crime Rate Change")
+rate.crime.vs.pop
+```
+
+``` {r echo=FALSE, message=FALSE, warnings=FALSE}
+rate.crime.vs.pop <- ggplot(tmp, aes(x=gdp_rate_change_value, y=total_crime_rate_change_value)) +
+                  geom_point(shape=1) +
+                  stat_smooth(method = "lm", col = "red") +
+                  ggtitle("GDP Rate Change and Total Crime Rate Change") +
+                  labs(x ="GDP Rate Change", y="Total Crime Rate Change")
+rate.crime.vs.pop
+```
+
+This is where the data got very interesting.
+As we can see there is a very signifigant change on the Difference in the rate of crime based on the population in a negative fashion.
+This means that every time the population difference increase the rate of crime decreases.
+But on the other side of thins the GDP change actually increases the Crime change.
+Both of these factors were statistically signifigant, meaning the both affect the difference in the crime rate.
+
+### What this all means.

### 2017-03-08 resource/index.Rmd
finished conclusion

@@
 ### What this all means.
+
+In conclusion we have found that in terms of actual *Crime Rate* population and GDP have very little affect on it and no affect in a statistically signifigant way.
+And while there are trends that can be seen in the data when looking at the graphs it is still not to a great enought degree to consider those two variables as major drivers in the *Crime Rate*.
+What is interesting is when we start to look at the differences between the many different values.
+After running our models we have found that the change in the Population difference has a very pronounced affect on the crime difference.
+So as the population difference increases the crime difference decreases.
+The GPD difference though has an opposite affect on crime diffence.
+As the GDP difference increases the crime difference decreases.
+This seems to go against what can be thought of as common held beliefs.
+The idead that when you live in a larger city there are more crimes, is true yet a person actual likelihood of committing a crime is decreased.
+And where one would think that the better a cities economy is the less crime we foud the opposite.
+All this points to is that there is much more to why crime happens then just the population and Gross Domestic Product.
+While these two things can have an affect on it there is still much research that needs to be done in order to truly understand all of the nuances that go into this.
+
+_If you want to look at our working code you can go to the repo [here](https://github.com/czross/info370-final)_

### 2017-03-08 resource/index.Rmd
pulling

--- 2017-03-08 resource/index.Rmd
@@
 ---
+<br>
 
@@
 library(rmarkdown)
+library(pander)
 source("resources.R")
 ```
+<br>
 
@@
 
-The two datasets that we used in this anaylsis are one for crime based data on U.S.
-cities over a 5 year time period (2010-2015) and one for the GDP for metropolitan areas in the U.S.
+For this research project we decided to use two datasets for our analysis; one for crime based data on U.S.
+cities over a 5 year time period (2010-2015), and one for the GDP for metropolitan areas in the U.S.
 over a 5 year time period (2010-2015).
@@
 
-Below is an example of what the crime dataset looks like (shows all 5 years):
-
-```{r, echo=FALSE}
-  kable(cities.all[1:5,])
-```
+You can see a sample of what the crime dataset looks like (shows all 5 years) [Here](../resource/crime-dataset.html).
 
@@
   select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015") %>%
-  kable()
+  pander()
 ```
@@
 <br> 
-
-###Crime Rate with GDP and Population
-
 So what we did is we took the newly made variable *Crime Rate* and compared it with the population of the different cities and their Gross Domestic Product per Person.
@@
 Theses numbers seem incredibly small but when you are talking about millions of dollars in GDP and thousands of people it becomes increasingly important.
-After computing this equation it was still found to not be statistically signifigant, meaning that the GDP per Person and the Population had no real impact on the Crime Rate.
-
-### Crime Rate Change
-
-After looking at just the crime rate over the 5 year time period we created a new variable which was the difference in crime rate, GDP per person, and Population change between the years.
-At this point we had started to work more with countinous data than count data, so we decided to use a standard linear regression to find the equation and signifigance of the data.
-
-``` {r echo=FALSE, message=FALSE, warnings=FALSE}
-rate.crime.vs.pop <- ggplot(tmp, aes(x=population_change_value, y=total_crime_rate_change_value)) +
-                  geom_point(shape=1) +
-                  stat_smooth(method = "lm", col = "red") +
-                  ggtitle("Population and Total Crime Rate Change") +
-                  labs(x ="Population", y="Total Crime Rate Change")
-rate.crime.vs.pop
-```
-
-``` {r echo=FALSE, message=FALSE, warnings=FALSE}
-rate.crime.vs.pop <- ggplot(tmp, aes(x=gdp_rate_change_value, y=total_crime_rate_change_value)) +
-                  geom_point(shape=1) +
-                  stat_smooth(method = "lm", col = "red") +
-                  ggtitle("GDP Rate Change and Total Crime Rate Change") +
-                  labs(x ="GDP Rate Change", y="Total Crime Rate Change")
-rate.crime.vs.pop
-```
-
-This is where the data got very interesting.
-As we can see there is a very signifigant change on the Difference in the rate of crime based on the population in a negative fashion.
-This means that every time the population difference increase the rate of crime decreases.
-But on the other side of thins the GDP change actually increases the Crime change.
-Both of these factors were statistically signifigant, meaning the both affect the difference in the crime rate.
-
-### What this all means.

### 2017-03-08 paper/paper.Rmd
working on formatting

@@
 ---
-title: "Final Project Paper"
 output: 
-  html_document: default
-  html_notebook: default
+  pdf_document:
+    citation_package: natbib
+    keep_tex: true
+    fig_caption: true
+    latex_engine: pdflatex
+    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-ms.tex
+title: "A review of Crime data and how GDP affects Crime Rates"
+author:
+- name: Kai Yoshino
+  affiliation: Univeristy of Washington
+- name: Isabella Spaletta 
+  affiliation: University of Washington
+- name: Rembrandt Q. Einstein
+  affiliation: Springfield University
+abstract: "This document provides an introduction to R Markdown, argues for its..."
+keywords: "pandoc, r markdown, knitr"
+date: "`r format(Sys.time(), '%B %d, %Y')`"
+geometry: margin=1in
+fontfamily: mathpazo
+fontsize: 11pt
+# spacing: double
+bibliography: ~/Dropbox/master.bib
+biblio-style: apsr
 ---

### 2017-03-08 resource/index.Rmd
clean up

@@
 ---
-<br>
 
@@
 
-For this research project we decided to use two datasets for our analysis; one for crime based data on U.S.
-cities over a 5 year time period (2010-2015), and one for the GDP for metropolitan areas in the U.S.
+For this research project we decided to use two datasets for our analysis; one for [crime based data](https://ucr.fbi.gov/crime-in-the-u.s/2015/crime-in-the-u.s.-2015/tables/table-8/table_8_offenses_known_to_law_enforcement_by_state_by_city_2015.xls/view) on U.S.
+cities over a 5 year time period (2010-2015), and one for the [GDP for metropolitan areas](https://www.bea.gov/national/index.htm) in the U.S.
 over a 5 year time period (2010-2015).
@@
 
-You can see a sample of what the crime dataset looks like (shows all 5 years) [Here](../resource/crime-dataset.html).
+You can see a sample of how we formatte the crime dataset (all 5 years) [Here](../resource/crime-dataset.html).
 
@@
 ```
+
+<br>
 
@@
 
+<br>
+
 ### Crime Rate Change
@@
 
+<br>
+
 ### What this all means.
@@
 _If you want to look at our working code you can go to the repo [here](https://github.com/czross/info370-final)_
+
+<br>
+<br>
+<br>

### 2017-03-08 paper/paper.Rmd
add tex file

--- 2017-03-08 paper/paper.Rmd
@@
 ---
-title: "Final Project Paper"
 output: 
-  html_document: default
-  html_notebook: default
+  pdf_document:
+    citation_package: natbib
+    keep_tex: true
+    fig_caption: true
+    latex_engine: pdflatex
+    template: ~\Documents\INFO370\info370-final\paper\svm-latex-ms.tex
+title: "A review of Crime data and how GDP affects Crime Rates"
+author:
+- name: Kai Yoshino
+  affiliation: Univeristy of Washington
+- name: Isabella Spaletta 
+  affiliation: University of Washington
+- name: Chrisopther Ross
+  affiliation: University of Washington
+-
+abstract: "The relationship between crime and the economy is complex.
+By analyzing the connection between crimes committed, GDP, and populations of metropolitan areas from the years 2010 to 2015, we seek to gain a deeper understanding on the driving forces behind these societal elements.
+The key findings of our research revealed that with a Poisson regression, there's no significant correlation between the count of total crimes committed per person (crime rate) in a metropolitan area and the predictors of that area's GDP per person or population.
+However, analyzing the rate changes in each variable via linear regression shows the opposite: a significant correlation.
+Looking at changes in both property and violent crime rates, we also found a significant correlation to both GDP rate changes and population changes.
+From this, our contribution to criminology and economics demonstrates that analyzing a metropolitan area's economy and population only scratches the surface for determining crime counts, suggesting that focus should be veered towards analyzing the rate changes in each variable.
+"
+keywords: "Crime, GDP, Poisson Regression, Linear Regression"
+date: "`r format(Sys.time(), '%B %d, %Y')`"
+geometry: margin=1in
+fontfamily: mathpazo
+fontsize: 11pt
+# spacing: double
+#bibliography: ~/Dropbox/master.bib
+#biblio-style: apsr
 ---
@@
 ********** Marisa
-The key findings of our research showed that a poisson regression showed no correlation between the GDP per person and crime per person in each metropolitan area with population as another independent variable.
-Between all of the years from 2010 to 2015, there was no significance between GDP per person, crime per person, and population of a metropolitan area for any compared years.
-From this, our contribution to society and information in general is that just looking at a metropolitan area's economy and population is not a good determining factor on the number of crimes permitted by person.
 

### 2017-03-08 paper/paper.Rmd
made pdf document

@@
 ---
-output: 
-  pdf_document:
-    citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-    template: ~\Documents\INFO370\info370-final\paper\svm-latex-ms.tex
+output: pdf_document
+  #pdf_document:
+    #citation_package: natbib
+    #keep_tex: true
+    #fig_caption: true
+    #latex_engine: pdflatex
+    #template: ~\INFO370\svm-r-markdown-templates\svm-latex-anon-ms
 title: "A review of Crime data and how GDP affects Crime Rates"
 author:
-- name: Kai Yoshino
-  affiliation: Univeristy of Washington
-- name: Isabella Spaletta 
-  affiliation: University of Washington
-- name: Chrisopther Ross
-  affiliation: University of Washington
--
+- Kai Yoshino, University of Washington
+- Isabella Spaletta, University of Washington 
+- Chrisopther Ross, University of Washington
+- Marisa Nanaku, University of Washington
+
 abstract: "The relationship between crime and the economy is complex.
@@
 
-#### Isabella Spaletta, Marisa Nanakul, Christopher Ross, Kai Yoshino
-##### Affiliation: University of Washington - Information School
-
-### Abstract
-
-********** Marisa
-
 ### Introduction
-
 Currently, there are many different sites and organizations that collect data on cities or large metropolitan areas.
@@
 Some of the related work we will discuss takes the change in economy after the 2008 financial crisis into account.
-One other research group took into account economic uncertainty whereas we will focus more on the count data of a metropolitan area (population, GDP).
-We want to look at the GDP and crime relationship affected by population to see if GDP per person and population of a metropolitan area have any sort of impact on the crimes per person committed within that metropolitan area.
+One other research group took into account economic uncertainty, whereas we will focus more on the count data of a metropolitan area (population, GDP, crime rate).
+We want to look at the GDP and population relationship affected by crime to see if GDP per person and population of a metropolitan area have any impact on the crime rate within that metropolitan area.
+
+Our analysis focuses on crime rate impact on city GDPs with population as another independent variable.
+The analysis is a unique contribution because rather than analyzing from the state level, we are choosing to focus on the metropolitan area level, adding an extra area of granularity.
+In terms of covariates, we are accounting for the population change between years along with the changes in GDP.
+This differentiates our work compared to previously produced work focusing on the variables by themselves.
+We are observing and comparing violent and property crime trends to metropolitan area GDP and population trends.
 
@@
 
+### Hypothesis
+The research question we seek to answer is: How do GDP and population affect property, violent, and total crime rates within U.S. metropolitan areas?
+
+We hypothesize that as population grows, crime rates would increase and that as GDP increases, the crime rate would decrease.
+We thought this because, generally speaking, one would think that with more people concentrated in an area the likelihood of crime would increase because there is less to control.
+Also, if the economy was doing well crime would potentially be decreasing since the economy would be in a state where people do not need to resort to committing crimes to survive.
+We also hypothesize that GDP decreases property crimes will increase, as populations will turn to crime to fix their economic troubles.
+
 ### Related Work
@@
 and countries around the world.
-In “Falling Together” [1], data from the U.S.
-Department of Justice from 2007-2010 was used to analyze the relationship between property crime rates and GDP per person in all 50 states.
+In "Falling Together" [1], data from the U.S.
+Department of Justice from 2007-2010 was used to analyze the relationship between property crime rates and GDP per person in all 50 U.S.
+states.
 The results found that property crime rates and GDP both decreased during this time frame.
@@
 
-“The Puzzling Relationship Between Crime and the Economy” [2] delves into the different conclusions economists and criminologists conclude about the relationship between crime and the economy.
+"The Puzzling Relationship Between Crime and the Economy" [2] delves into the clashing conclusions economists and criminologists draw from the relationship between crime and the economy.
 Criminologists believe that tough economic times increase the likelihood of a person committing a crime.
@@
 Economists believe better economic conditions increase the likelihood of a person committing a crime.
-With technology becoming more visible and ingrained into people’s lives, criminals are able to see what’s possible to steal.
+With technology becoming more visible and ingrained into people's lives, criminals are able to see what's possible to steal.
 Analysis on violent and property crimes from 1960-2012 suggests that there is no relationship between GDP and crime, and that crime is episodic.
 
-In “Economic growth and crime: does uncertainty matter?” [3] 25 countries (including the U.S.) are analyzed with a macroeconomic uncertainty metric to observe whether there is a relationship between crime and GDP.
+In "Economic growth and crime: does uncertainty matter?" [3] 25 countries (including the U.S.) are analyzed with a macroeconomic uncertainty metric to observe whether there is a relationship between crime and GDP.
 The researchers found that if GDP uncertainty is negative, the crime uncertainty is positive, meaning that crime uncertainty negatively impacts GDP uncertainty.
@@
 This is important because to keep the GDP accurate, we had to keep the cities combined as a metropolitan area, but we could account for the population and crimes by adding the counts together and doing analysis from the new totals.
-We also removed any metropolitan area from the GDP dataset that included multiple states and had to change any value in the data set that was NA to 0 to be able to combine these datasets properly.
+We also removed any metropolitan area from the GDP dataset that included multiple states and had to change any value in the dataset that was "NA" to "0" to be able to combine these datasets properly and include as many cities as we could in the analysis.
+Although "0" values may affect and weigh down the results, this was the best way for us to both keep as many cities as we could in the dataset and make the values calculable.
 
@@
 This came out to include 366 cities or metropolitan areas.
-We also added columns such as crime per person in each metropolitan area and gdp per person in each metropolitan area to run our analyses on.
+We also added columns such as "crime per person" and "gdp per person" in each metropolitan area in order to support our analysis.
 
-One thing to note is that to reproduce this research you will have to use the CSV we created or any other researcher could find other data sources for GDP and crime for the years 2010 to 2015 and create their own new dataset.
-You must also change any NA in the datset to 0 to combine the data.
+One thing to note is that to reproduce this research you will have to use the CSV we created.
+Alternatively, any other researcher could find other data sources for GDP and crime for the years 2010 to 2015 and create their own new dataset.
+You must also change any "NA" values in the dataset to "0" to combine the data.
 
-Another thing to note is that 
+Some definitions to note are 
 'violent crimes' is defined by four offenses: murder and nonnegligent manslaughter, rape, robbery, and aggravated assault. 
 'property crimes' is defined by four offenses: burglary, larceny theft, motor vehicle theft, and arson.
+Crime rate, as defined by this research, is the total number of crimes in a metropolitan area divided by the population of that metropolitan area to become crime per person.
 
-### Analysis
+### Methods and Analysis 
+In order to research this dataset we first chose to take a look at the crime rate data via Poisson regression using population and GDP per person as outcomes.
+Our crime rate data was measured by dividing the total counts of those crimes by metropolitan area and dividing them by that area's population.
+We analyzed crime using three different outcomes for our regressions: violent crime, property crime, and total crime (violent and property combined.) For covariates, we used GDP per person and population by metropolitan area.
+We reorganized the structure of the data from wide format to long format - placing the variables for each year in the same column instead of separate columns for each year.
+This allowed us to try to predict outcomes for all years in our dataset instead of one year at a time as the data was originally structured.
+Violent crime rates were usually lower than property crime rates.
+We ran into issues where some cities had not fully reported crimes, putting null values in our dataset.
+These occurrences were low enough that we decided to set them to zero so we would not have to throw out data for entire cities.
+After seeing the non-significant results from our Poisson regression for total crime we decided to take a further look at the data.
+While total occurrences of our variables might not show much correlation, there was a possibility that changes in our hypothesized crime predictors could significantly relate to changes in our crime outcomes.
+So we decided to switch gears and calculate the rate changes in both our covariates and our outcomes to see if there was a correlation.
+We decided that a linear regression was more appropriate for analyzing this data as we were no longer using count data.
+To do this we simply took the difference of crime rates between each consecutive year for property, violent, and total crime rates.
+We used the difference in GDP per person for consecutive years and the rate change in population for consecutive years as our covariates.
+Similarly to our first analysis, we structured our data in long format with columns that included each consecutive years pair for each variable.
 
 ### Results
+For our Poisson regressions, we found no significant correlations between crime rate and our covariates of population and GDP per person.
+This was the case for all outcomes of property, violent, and total crime.
+
+![poisson](poisson.png)
+
+After viewing the extremely low confidence intervals of our regression for total crime we decided to reframe our data and use a linear regression on the crime rate change instead of just the crime rate.
+For our linear regression we found significant correlations between crime rate changes of all types and our covariates of changes in population (negatively) and changes in GDP per person (positively).
+
+![linear](linear.png)
 
 ### Discussion
+We initially hypothesized that there would be a correlation between counts of GDP per person, population and crime per person.
+However our analysis found that this was not the case.
+It is likely that there are far too many factors involved in predicting crime rates, and that population and GDP alone, cannot confidently do the job.
+Possible confounds could be education, unemployment, wealth disparity, unreported population or crime, and likely many more.
+
+The results of our linear regression on rate changes in our variables were also quite surprising.
+We hypothesized that higher population and lower GDP would result in higher crime rates.
+But our regressions shows that as population decreases, crime rates of all kinds increase, and as GDP increases, crime rates of all kind increase.
+Although our hypothesis referred to total variable counts, we believe it extends to rate changes in our variables similarly.
+However rate changes showed the opposite of what we hypothesized.
+Additionally we hypothesized that low GDP would correlate to high property crime at a lower coefficient than to violent crime, as more people would be committing crimes to acquire more money in a poor economy.
+However, we found in our linear regression that when GDP increases, violent crimes increase at a greater rate than property crimes.
+So GDP seems to affect violent crimes more than property crimes, which is not what we expected.
+
+Because our results are off from our original hypothesis, we are naturally skeptical about them.
+As stated earlier, it appears predicting crime rates is not as cut and dry as looking at GDP and population.
+It is likely that crime rate changes are also influenced by many other outside factors.
+There are numerous possible next steps to take with these results.
+Future work could greatly improve, reinforce, or reinterpret our results and better explain how crime rates can be predicted.
 
@@
 While we looked at GDP and crime rate, to extend the work we have done here, other researchers can also compare GDP and crime rate per city with other independent variables other than population that may also affect the correlation.
-One example is other related work looks at economic uncertainty as an independten variable and there are other covariates that could possibly show a correlation between the GDP and crime rate of a city such as education, population density, health metrics, and many more.
+One example is other related work uses economic uncertainty as an independent variable and there are other variables that could possibly show a correlation between the GDP and crime rate of a city such as education, population density, health metrics, and many more.
 
 One other aspect to consider is the particular datasets used and how the data is measured.
-Since we are doing by metropolitan area, crime rate to us is defined as total number of crimes divided by the population of each city or metropolitan area within the dataset.
-The GDP dataset that we chose to use was in a very unfortunate format when trying to combine it with the crime dataset, so I would suggest that in future work an alternative GDP dataset is used that is by actual city in the US rather than metropolitan area.
+Since we are sorting by metropolitan area, crime rate to us is defined as total number of crimes divided by the population of each city or metropolitan area within the dataset.
+The GDP dataset that we chose to use was in a very unfortunate format when trying to combine it with the crime dataset, so we would suggest that in future work, an alternative GDP dataset defined by U.S.
+city rather than by metropolitan area be used.
 
@@
 
-[3] Goulas, Eleftherios, and Athina Zervoyianni. "Economic Growth and Crime: Does Uncertainty Matter?" Applied Economics Letters 20.5 (2013): 420-27. Web.
+[1] The Economist Online. "Falling Together." The Economist. The Economist Newspaper, 26 Sept. 2011. Web. 05 Mar. 2017.
 
@@
 
-[1] The Economist Online. "Falling Together." The Economist. The Economist Newspaper, 26 Sept. 2011. Web. 05 Mar. 2017.
+[3] Goulas, Eleftherios, and Athina Zervoyianni. "Economic Growth and Crime: Does Uncertainty Matter?" Applied Economics Letters 20.5 (2013): 420-27. Web.

### 2017-03-08 resource/resources.R

@@ -228,4 +228,3 @@ linear.pop.gdp.property.crime <- lm(property_crime_rate_change_value ~ populatio
 
 plot(lm(total_crime_rate_change_value ~ population_change_value, data = linear.change.data))
 summary(linear.pop.gdp.property.crime)
-

### 2017-03-08 paper/paper.Rmd
finished

@@
     #template: ~\INFO370\svm-r-markdown-templates\svm-latex-anon-ms
-title: "A review of Crime data and how GDP affects Crime Rates"
+title: "A Review of Crime Data and how GDP and Population Affects Crime Rates"
 author:
